Relative Stable Maps in Genus One via Central Alignments

### 2018-03-21 genus-one-relative-maps.tex

\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\setcounter{tocdepth}{2}

@@

\begin{remark}[Explaining compatibility]
First I wish to define the first line $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$; when looking at a geometric point $\Spec(k=\bar k)$ it is well-known that \[ Q_{\rm{min}}^{\rm{prestable}}=\prod_{q\ \text{node}}\mathbb N_q.\]
Let me assume that all the nodes are internal, i.e. map to $Y$, for all the action happens there, and also that there is only one internal connected component $Z$, which has arithmetic genus one. Then I define $Q_{\rm{min}}^{\rm{log.map}}$ as the saturation of the quotient of
\begin{equation}\label{eq:logminmod} \prod_{q\ \text{node}}\mathbb N_q \times \prod_{S\ \text{internal irr. compo.}}\mathbb N_S\end{equation}
by the relation generated in the groupification of the latter by
\[\langle \rm{Rel}_q=(u_q,\ldots,1,\ldots,-1,\ldots),\ q\ \text{internal node} \rangle\]
where $1$ is in position $S_q^+$ (adjacent to $q$ and further from the circuit), $-1$ in position $S_q^-$ (adjacent to $q$ and closer to the circuit), and $u_q$ in position $q$ is defined as to include the tropical balancing condition (I don't know whether this is standard):
\[u_q=f_*[T_q]\cdot [Y]+\sum_{R_i \text{external compo. attached to }T_q} m^{(i)}- \sum_{x_j\in T_q}\alpha_j,\]
where $T_q$ is the tree of internal components that $q$ separates from the circuit. Notice that the nodes separating $Z$ from the external components have no $1$, so the corresponding relations look like \[(u_q,\ldots,-1,\ldots).\]
The map $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$ is induced by the inclusion of the node factors in \eqref{eq:logminmod}. If we call $e_q$ the image in $Q_{\rm{min}}^{\rm{log.map}}$ of the $q$-th basis vector, then we can verify that the following equations hold: chosen any internal component $S_i$ and any two external component $R_i^1, R_i^2$
such that their path to the circuit goes through $S_i$, then
\begin{equation}
 \sum_{q\in[R_i^1,S_i]}u_qe_q=\sum_{q\in[R_i^2,S_i]}u_qe_q.
\end{equation}
\textcolor{red}{Question: is this monoid factorisation condition a closed one?}
\end{remark}

\begin{remark}
 Here is a tidy example: start with an elliptic curve $E$ contracted to $Y$, with two external components $R_1, R_2$ attached to it and a single marking $x\in E$. The tropical picture would be:
 \begin{figure}
 \begin{tikzpicture}[radius=2pt,scale=0.5]
 \draw[fill=black] (0,0) circle;
 \draw (0,0) -- (4,0);
 \draw[fill=black] (0,2) circle;
 \draw[fill=black] (0,4) circle;
 \draw (0,2) -- (2,3);
 \draw (0,4) -- (2,3);
 \draw (2,3) -- (4,3);
 \draw[fill=white] (2,3) circle;
 \draw[->] (2,1.5) -- (2,.5);
 \end{tikzpicture}
 \end{figure}
The dual picture of the compatibility condition would be:
\begin{tikzpicture}[scale=0.8]
 \draw (0,0) -- (0,2)  (0,0)--(2,0);
 \draw[->] (1,-1.5) -- (1,-.5);
 \draw (0,-4) -- (0,-2) (0,-4) -- (2,-4) (0,-4) -- (2,-2) node[above right]{$e_1=e_2$};
 \draw (1,-2) node[left]{$e_2>e_1$} (2,-3) node[right]{$e_1>e_2$};
 \draw[->] (2.5,1)--(3.5,1);
 \draw[gray!=30] (4,0) -- (4,2)  (4,0)--(6,0);
 \draw (4,0) -- (4.6,2) node[above right]{$m_1/m_2$};
\end{tikzpicture}
which is compatible with the fact that $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$ is given by:
$e_1\mapsto m_2, e_2\mapsto m_1$.
\end{remark}

\section{The relative space is equal to the closure of the nice locus for $(\PP^N,H)$}
\noindent In general we do not know very much about the space we have just defined. The aim of this section is to show that, in the case where $X=\PP^N$ and $Y=H$ is a hyperplane, the space is proper and irreducible of the expected dimension. As such it has a fundamental class, which we can use to define reduced relative Gromov--Witten invariants in genus one.

@@

\begin{lem}\label{Nice locus is irreducible}
The nice locus $\VZc{1}{\alpha}{\PP^N|H}{d}^{\circ}$ is irreducible.
\end{lem}
\begin{proof}
By definition the contraction radius $\delta$ is compatible with the map, i.e. the subcurve $C_0\subseteq C$ where $\lambda<\delta$ is equal to the maximal connected genus one subcurve contracted by $f$. Hence when the source curve is irreducible we must have $\delta=0$, and we see that the central alignment on $C$ is uniquely and trivially determined. Thus to specify a point in the nice locus we only need to specify the source curve $C$ (as a scheme) and the map $f$. A parametrisation can be given from the vector bundle:\marginnote{I'm sure there's a simpler way to say this - Navid}
\[ \operatorname{Vb}\left(\pi_*\OO_{\mathcal E}(\sum_{j=n+1}^{n+\delta}\sigma_j)\oplus\pi_*\OO_{\mathcal E}(\sum_{j=1}^{n+\delta}\sigma_j)^{\oplus r}\right) \quad \text{on} \quad \mathcal{M}_{1,n+\delta}\]
where $\pi\colon\mathcal E\to\mathcal M$ is the universal curve and $\delta=d-\sum\alpha_i$.
\end{proof}

\subsection{Justifying the novel condition}
Remember that we are trying to impose conditions which determine the closure of the nice locus inside the moduli space of centrally aligned maps. Here is one example where we show by a dimensional computation that the novel condition must be included if we hope to determine the closure of the nice locus.
\begin{example}
Consider $\M{1}{(3)}{\PP^1|H}{3}$. This moduli space has virtual dimension $6+1-3=4$. Here is a parametrisation of the nice locus: choose an object $(E,p) \in \mathcal{M}_{1,1}$, and let $s_0$ be the natural section $s_0 \colon\OO_E\hookrightarrow\OO_E(3p)$ and $s_1$ any other section of $\OO_E(3p)$ not vanishing at $p$ (notice that $\h^0(E,\OO_E(3p))=3$). Then $(E,p,[\lambda s_0,s_1])$ gives a well-defined element of the nice locus for $\lambda \neq 0$.

@@

\begin{remark}
 For the non-expert reader, it might be convenient to recall that the extra data of a centrally aligned log structure on the curves gives a modular interpretation to the iterated blow-up procedure of Vakil and Zinger. In particular the novel condition on the log structure can be translated into a closed condition on the exceptional loci of these blow-ups. 
\end{remark}

@@

It thus remains to show that, given a relative radially aligned map, we can smooth it to one in the nice locus. \textcolor{red}{This is done by considering different cases locally, then gluing. \marginpar{What do we exactly mean by this? How does gluing work in the centrally aligned setting?}}

@@

\section{Novel Condition 2.0}

\section{Basic facts about centrally aligned curves}
\begin{prop}[{\cite[Proposition 4.6.2.2]{RSPW}}]
The morphism $\MM^{\rm{ctr}}_{1,n}\to\MM^{\dagger}_{1,n}$ is a log-modification.
\end{prop}

>>>>

\newcommand{\TT}{\operatorname{T}}
\newcommand{\oM}{\overline{\mathcal{M}}}
\newcommand{\M}[4]{\overline{\mathcal{M}}_{#1,#2}(#3,#4)}
\newcommand{\Q}[4]{\mathcal{Q}_{#1,#2}(#3,#4)}
\newcommand{\Qe}[4]{\mathcal{Q}^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\Qt}[4]{\widetilde{\mathcal Q}_{#1,#2}(#3,#4)}
\newcommand{\QG}[4]{\mathcal{Q}G_{#1,#2}(#3,#4)}
\newcommand{\QGe}[4]{\mathcal{Q}G^{\epsilon}_{#1,#2}(#3,#4)}
\newcommand{\D}[3]{\mathcal{D^Q}(#1,#2,#3)}
\newcommand{\E}[3]{\mathcal{E^Q}(#1,#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\tVZc}[4]{\widetilde{\mathcal{V\!Z}}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZc}[4]{\mathcal{V\!Z}^{\rm{ctr}}_{#1,#2}(#3,#4)}
\newcommand{\VZcLi}[4]{\mathcal{V\!Z}^{\rm{ctr, Li}}_{#1,#2}(#3,#4)}
\newcommand{\N}{\mathbb{N}}
\newcommand{\OO}{\mathcal{O}}
\renewcommand{\to}{\rightarrow}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\EE}{\mathbf{E}}
\renewcommand{\L}{\mathcal L}
\newcommand{\LL}{\mathbf{L}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\Aaff}{\mathbb{A}}
\newcommand{\kfield}{\Bbbk}
\newcommand{\comp}{\chi}
\newcommand{\sst}{\sigma^{\operatorname{ss}}}
\newcommand{\Pic}{\operatorname{Pic}}
\newcommand{\Def}{\operatorname{Def}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Proj}{\operatorname{Proj}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Gm}{\mathbb{G}_{\text{m}}}
\newcommand{\virt}[1]{[#1]^{\operatorname{virt}}}
\newcommand{\vip}[1]{[#1]^{\operatorname{prod}}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\HH}{\operatorname{H}}
\newcommand{\Achow}{\operatorname{A}}
\newcommand{\pt}{\operatorname{pt}}
\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\ba}{\begin{aligned}}
\newcommand{\ea}{\end{aligned}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bsm}{\left(\begin{smallmatrix}}
\newcommand{\esm}{\end{smallmatrix}\right)}                   
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}
\newcommand{\barr}{\begin{displaymath}\begin{array}{cccc}}
\newcommand{\earr}{\end{array}\end{displaymath}}
\newcommand{\barrl}{\begin{displaymath}\begin{array}{lcl}}
\newcommand{\earrl}{\end{array}\end{displaymath}}
\newcommand{\barl}{\begin{displaymath}\begin{array}{l}}
\newcommand{\earl}{\end{array}\end{displaymath}}
\newcommand{\bxym}{ \begin{displaymath}\xymatrix }
\newcommand{\exym}{\end{displaymath}}
\newcommand{\bcd}{\begin{center}\begin{tikzcd}}
\newcommand{\ecd}{\end{tikzcd}\end{center}}
\newcommand{\R}{\operatorname{R}^{\bullet}}
\newcommand{\dvr}{\Delta}
%\newcommand{\sslash}{\mathbin{/\mkern-6mu/}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\Isom}{\text{Isom}}
\newcommand{\pr}{\operatorname{pr}}
\newcommand{\ev}{\operatorname{ev}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\vdim}{\operatorname{vdim}}
\newcommand{\ildef}[1]{\emph{#1}}
\newcommand{\om}[1]{\mathcal{#1}}
\newcommand{\h}{\operatorname{h}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\RR}{\textbf{R}}
\newcommand{\NN}{\operatorname{N}}
\newcommand{\ovm}[1]{\overline{\mathcal{#1}}}
\newcommand{\ovt}[1]{\widetilde{\mathcal{#1}}}
\newcommand{\ov}[1]{\overline{#1}}

@@

\setcounter{tocdepth}{1}

@@

\begin{remark}[Explaining compatibility]
First we need to define the monoids and maps involved. We start with $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$; when looking at a geometric point $\Spec(k=\bar k)$ it is well-known that \[ Q_{\rm{min}}^{\rm{prestable}}=\prod_{q\ \text{node}}\mathbb N_q.\]
Let us assume that all the nodes are internal, i.e. map to $Y$, for all the action happens there, and also that there is only one internal connected component $Z$, which is a curve of arithmetic genus one. Then we define $Q_{\rm{min}}^{\rm{log.map}}$ as the saturation of the quotient of
\begin{equation}\label{eq:logminmod} \prod_{q\ \text{node}}\mathbb N_q \times \prod_{S\ \text{internal irr. compo.}}\mathbb N_S\end{equation}
by the relation generated in the groupification of the latter by
\[\langle \rm{Rel}_q=(u_q,\ldots,1,\ldots,-1,\ldots),\ q\ \text{internal node} \rangle\]
where $1$ is in position $S_q^+$ (adjacent to $q$ and further from the circuit), $-1$ in position $S_q^-$ (adjacent to $q$ and closer to the circuit), and $u_q$ in position $q$ is defined as to include the tropical balancing condition (I don't know whether this is standard):
\[u_q=f_*[T_q]\cdot [Y]+\sum_{R_i \text{external compo. attached to }T_q} m^{(i)}- \sum_{x_j\in T_q}\alpha_j,\]
where $T_q$ is the tree of internal components that $q$ separates from the circuit. Notice that the nodes between $Z$ and an external component $R_i$ have no $1$, so the corresponding relations look like \[(u_q,\ldots,-1,\ldots).\]
The map $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$ is induced by the inclusion of the node factors in \eqref{eq:logminmod}. If we call $e_q$ the image in $Q_{\rm{min}}^{\rm{log.map}}$ of the $q$-th basis vector, then we can verify that the following equations hold: chosen any internal component $S_i$ and any two external component $R_i^1, R_i^2$
such that their path to the circuit goes through $S_i$, then
\begin{equation}
 \sum_{q\in[R_i^1,S_i]}u_qe_q=\sum_{q\in[R_i^2,S_i]}u_qe_q.
\end{equation}
\textcolor{red}{Question: is this monoid factorisation condition a closed one?}
\end{remark}

\begin{remark}
 Here is a tidy example: start with an elliptic curve $E$ contracted to $Y$, with two external components $R_1, R_2$ attached to it and a single marking $x\in E$. The tropical picture would be:
\begin{center}
 \begin{tikzpicture}[radius=2pt,scale=0.5]
 \draw[fill=black] (0,0) circle;
 \draw (0,0) -- (4,0);
 \draw[fill=black] (0,2) circle;
 \draw[fill=black] (0,4) circle;
 \draw (0,2) -- (2,3);
 \draw (0,4) -- (2,3);
 \draw (2,3) -- (4,3);
 \draw[fill=white] (2,3) circle;
 \draw[->] (2,1.5) -- (2,.5);
 \end{tikzpicture}
\end{center}
The dual monoids in the compatibility condition give us the following nice picture:
\begin{center}
\begin{tikzpicture}[scale=0.8]
 \draw (0,0) -- (0,2)  (0,0)--(2,0);
 \draw[->] (1,-1.3) -- (1,-.5);
 \fill[gray!30] (0,-4) -- (0,-2) -- (2,-2) -- (0,-4);
 \fill[gray!50] (0,-4) -- (2,-4) -- (2,-2) -- (0,-4);
 \draw (0,-4) -- (0,-2) (0,-4) -- (2,-4) (0,-4) -- (2,-2) node[above right]{$e_1=e_2$};
 \draw (1,-2) node[above]{$e_2>e_1$} (2,-3) node[right]{$e_1>e_2$};
 \draw[<-] (2.5,1)--(3.5,1);
 \draw[gray!=30] (4,0) -- (4,2)  (4,0)--(6,0);
 \draw (4,0) -- (4.6,2) node[above right]{$m_1/m_2$};
 \draw[<-,dashed] (2.7,-1.3)--(3.5,-.5);
\end{tikzpicture}
\end{center}
So, for example, the factorisation exists for the subcone $e_2>e_1$ only when $m_1>m_2$, which is compatible with the fact that $\mathbb N^2=Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}=\mathbb N$ is given by
$e_1\mapsto m_2, e_2\mapsto m_1$.
\end{remark}

\section{The relative space is the closure of the nice locus for $(\PP^N,H)$}
\noindent In general we do not know very much about the space we have just defined. The aim of this section is to show that, in the case where $X=\PP^N$ and $Y=H$ is a hyperplane, the space is proper and irreducible of the expected dimension. As such it has a fundamental class, which we can use to define reduced relative Gromov--Witten invariants in genus one.

@@

\begin{lem}\label{Nice locus is irreducible}
The nice locus $\VZc{1}{\alpha}{\PP^N|H}{d}^{\circ}$ is irreducible.
\end{lem}
\begin{proof}
By definition the contraction radius $\delta$ is compatible with the map, i.e. the subcurve $C_0\subseteq C$ where $\lambda<\delta$ is equal to the maximal connected genus one subcurve contracted by $f$. Hence when the source curve is irreducible we must have $\delta=0$, and we see that the central alignment on $C$ is uniquely and trivially determined. Thus to specify a point in the nice locus we only need to specify the source curve $C$ (as a scheme) and the map $f$. A parametrisation can be given from the vector bundle:
\[ \operatorname{Vb}\left(\pi_*\OO_{\mathcal E}(\sum_{j=n+1}^{n+\delta}\sigma_j)\oplus\pi_*\OO_{\mathcal E}(\sum_{j=1}^{n+\delta}\sigma_j)^{\oplus r}\right) \quad \text{on} \quad \mathcal{M}_{1,n+\delta}\]
where $\pi\colon\mathcal E\to\mathcal M$ is the universal curve and $\delta=d-\sum\alpha_i$.
\end{proof}

\subsection{Justifying the novel condition by dimensional reasons}
Here is a dimensional computation in an explicit example which shows that being in the closure of the nice locus must impose some condition on the log structure.
\begin{example}
Consider $\M{1}{(3)}{\PP^1|H}{3}$. This moduli space has virtual dimension $6+1-3=4$. Here is a parametrisation of the nice locus: choose an object $(E,p) \in \mathcal{M}_{1,1}$, and let $s_0$ be the natural section $s_0 \colon\OO_E\hookrightarrow\OO_E(3p)$ and $s_1$ any other section of $\OO_E(3p)$ not vanishing at $p$ (notice that $\h^0(E,\OO_E(3p))=3$). Then $(E,p,[\lambda s_0,s_1])$ gives a well-defined element of the nice locus for $\lambda \neq 0$.

@@

\begin{remark}
 For the reader who is unfamiliar with centrally aligned log curves, it might be convenient to recall that they provide a modular interpretation of the iterated blow-up procedure of Vakil and Zinger. In particular it is natural that not all of the exceptional divisors be contained in the closure of the nice locus for every $\alpha$; this geometric information is contained in the compatibility condition.
\end{remark}

\subsection{Proof sketch} \textcolor{red}{The compatibility condition should imply that the map admits a log-enhancement;} if we can prove that there is a smooth, connected and proper moduli space above, then we automatically get that the Gathmann-like relative space is closed and irreducible, and it obviously contains the nice locus; hence, if we can prove that the latter is dense, we win.

It thus remains to show that, given a relative centrally aligned map, we can smooth it to one in the nice locus. The only interesting case \textcolor{red}{should be} that the core is contracted to a point of $Y$. Here is how we prove it:
\begin{enumerate}
 \item we consider the core (say it's a smooth elliptic curve) marked with all the special points (markings and nodes); consider $E\times\dvr$ and blow it up in all the points of the central fiber that originally corresponded to nodes; we carefully and iteratively blow up smooth points of the central fiber in order to reconstruct the curve $C$ that we started with. In fact we may perform weighted instead of ordinary blow-ups, picking the ideal $(z,t^r)$ where $z$ is a vertical coordinate; this has the effect of introducing an $A_{r-1}$ surface singularity at the corresponding node of the central fiber. We do so by choosing a homomorphism $Q_{\rm{min}}^{\rm{log.map}}\to\mathbb N$ and composing it with $Q_{\rm{min}}^{\rm{prestable}}\to Q_{\rm{min}}^{\rm{log.map}}$ to determine the relevant $r$ for every node.
 \item Let $E,S_1,\ldots,S_k$ be the irreducible components of $f^{-1}(Y)$; we define a line bundle on the smoothing by
 \[\mathcal L=\OO_{\cC}\left(\sum \gamma_iS_i\right)\left(\sum \alpha_ix_i\right).\]
 Using intersection theory on the surface (recall that the local multiplicity of intersection of the two branches of the node underlying an $A_{r-1}$ singularity is $\frac{1}{r}$) we check that the compatibility condition is precisely what we need in order to be able to choose $\gamma_i$ adequately so that the following equations hold:
 \begin{align*}
  0&=\deg(\mathcal L_{|S_i}) \\
  m^{(i)}&=\deg(\mathcal L_{|R_i}) \\
  \mathcal O_E&\cong\mathcal L_{|E}
 \end{align*}
 the last from Gathmann's condition.
 \item It is not clear that sections of $\mathcal L$ are unobstructed. In order to extend them we pass to $\overline{\mathcal C}$.
\end{enumerate}

\section{Old}

\subsection{Basic facts about centrally aligned curves}
\begin{prop}[{\cite[Proposition 4.6.2.2]{RSPW}}]
The morphism $\MM^{\rm{ctr}}_{1,n}\to\MM^{\dagger}_{1,n}$ is a log-modification.
\end{prop}

\subsection{Moral justification of the novel condition}

@@

\subsection{Novel Condition 2.0}

\subsection{Details of the proof}

@@
